\assignment

### 2017-04-02 project/paper.rnw
Taking a break due to lazy load errors.

\newcommand\assignment{Stat 534 Project: \\
Extrapolation from Poisson Process Intensity Surface Models}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{\today}

@@

<<hw4_fitplot, fig.pos = 'p', fig.height = 8, fig.cap = 'Estimated log-intensity surface, log-scale prediction standard errors, and relative standard error from models fit using all events in the full simulation region (left) and events in a subset of the simulation region (right). The regions and event locations are overlaid in white.'>>=
par(mfrow = c(3, 2), mar = c(1, 0, 1.2, 2), las = 2, cex = 1)

@@

If the general structure of the site (such as the number and shape of the
high-intensity regions and the type of background process) is known then the
most appropriate model for the anomaly intensity could be a mixture model,
but in practice these details are rarely available. Therefore, practitioners
prefer flexible models without defined structure. Polynomial models are
conceptually simple, and capable of appriximating complicated surfaces. I
fit \(p\)-degree polynomial models of the form
\begin{equation*}
\log\left(\lambda\left(x, y\right)\right)
= \sum_{i = 0}^{p} \sum_{j=0}^{p-i} \beta_{ij} x^{i} y^{j}
\end{equation*}
for \(p = 2, 3, \dots, 12\). Here, the coordinates \(x\) and \(y\) are
standardized to have mean 0 and standard variance 1 in order to reduce
numerical instability for the larger exponents.

use \texttt{polynom()} not \texttt{poly()}!

<<polyfit1, cache = TRUE, dependson = 'sim2000obs', warning = FALSE, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = ''>>=
# Do a simple fit where spatstat sets up the quadrature so we can see where it
# puts the dummy points.
ex_fit1 <- ppm(ex_ppp ~ x + y)

# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
# TODO: Figure out weights.
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))

@@

par(mfrow = c(4, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, main = '2nd Degree')
plot(43560 * ex_p3$estimate, main = '3rd Degree')
plot(43560 * ex_p4$estimate, main = '4th Degree')
plot(43560 * ex_p5$estimate, main = '5th Degree')
plot(43560 * ex_p6$estimate, main = '6th Degree')
plot(43560 * ex_p7$estimate, main = '7th Degree')
plot(43560 * ex_p8$estimate, main = '8th Degree')
plot(43560 * ex_p9$estimate, main = '9th Degree')
plot(43560 * ex_p10$estimate, main = '10th Degree')
plot(43560 * ex_p11$estimate, main = '11th Degree')
plot(43560 * ex_p12$estimate, main = '12th Degree')
plot(43560 * density(ex_ppp_full) * area(ex_ppp_full) / area(ex_ppp),
     main = 'Kernel Density')
@

<<polyse1, cache = TRUE, dependson = 'polyfit1', fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = ''>>=
par(mfrow = c(4, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(ex_p2$se / ex_p2$estimate, main = '2nd Degree\nRSE')
plot(ex_p4$se / ex_p4$estimate, main = '4th Degree\nRSE')
plot(ex_p6$se / ex_p6$estimate, main = '6th Degree\nRSE')

@@

>>>>

\newcommand\assignment{Stat 534 Project: \\
Extrapolation from Poisson Process Intensity Surface Models}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{April 2, 2017}

@@

<<hw4_fitplot, fig.pos = 'p', fig.height = 8, fig.cap = 'Estimated log-intensity surface, log-scale prediction standard errors, and relative standard errors from models fit using all events in the full simulation region (left) and events in a subset of the simulation region (right). The regions and event locations are overlaid in white.'>>=
par(mfrow = c(3, 2), mar = c(1, 0, 1.2, 2), las = 2, cex = 1)

@@

If the general structure of the site (such as the number and shape of the
high-intensity regions and the type of background process) is known then the
most appropriate model for the anomaly intensity could be a mixture model or
a trend surface with a particular form, but in practice these details are
rarely available. Therefore, practitioners prefer flexible models without
defined structure. I fit \(p\)-degree polynomial models of the form
\begin{equation*}
\log\left(\lambda\left(x, y\right)\right)
= \sum_{i = 0}^{p} \sum_{j=0}^{p-i} \beta_{ij} x^{i} y^{j}
\end{equation*}
for \(p = 2, 3, \dots, 12\) using 500,000 dummy points randomly selected from
a uniform distribution over the observed region. The coordinates \(x\) and
\(y\) are standardized to have mean 0 and variance 1 in order to reduce
numerical instability for the larger exponents.

<<polyfit1, cache = TRUE, dependson = 'sim2000obs', warning = FALSE>>=
# Do a simple fit where spatstat sets up the quadrature so we can see where it
# puts the dummy points.
ex_fit1 <- ppm(ex_ppp ~ x + y)

# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))

@@

ex_kde <- density(ex_ppp_full)
@

Figure~\ref{fig:polyplot1} shows the predicted intensity surfaces for the
polynomial models, along with a kernel density estimate for comparison. The
2nd-order polynomial surface is too simple to capture the bimodal surface,

The kernel density estimate is computed using the \texttt{spatstat}'s default
``simple rule of thumb'' to select the smoothing bandwidth. Because the
anomalies located more than three feet from a transect were not observed, I
would expect the kernel density estimate to show ridges of high intensity
along the transects. As it turns out, the default bandwith is large enough to
smooth across multiple transects so the ridges do not appear. However, the
kernel density estimate is biased low because not all anomalies are observed;
I divide the estimate by the proportion of the area observed as an approximate
bias correction.

<<polyplot1, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = 'Predicted intensity surfaces for eleven polynomial models, and a kernel density estimate for comparision.'>>=
par(mfrow = c(4, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, zlim = c(0, 500), main = '2nd Degree')
plot(43560 * ex_p3$estimate, zlim = c(0, 500), main = '3rd Degree')
plot(43560 * ex_p4$estimate, zlim = c(0, 500), main = '4th Degree')
plot(43560 * ex_p5$estimate, zlim = c(0, 500), main = '5th Degree')
plot(43560 * ex_p6$estimate, zlim = c(0, 500), main = '6th Degree')
plot(43560 * ex_p7$estimate, zlim = c(0, 500), main = '7th Degree')
plot(43560 * ex_p8$estimate, zlim = c(0, 500), main = '8th Degree')
plot(43560 * ex_p9$estimate, zlim = c(0, 500), main = '9th Degree')
plot(43560 * ex_p10$estimate, zlim = c(0, 500), main = '10th Degree')
plot(43560 * ex_p11$estimate, zlim = c(0, 500), main = '11th Degree')
plot(43560 * ex_p12$estimate, zlim = c(0, 500), main = '12th Degree')
plot(43560 * ex_kde * area(ex_ppp_full) / area(ex_ppp), zlim = c(0, 500),
     main = 'Kernel Density')
@

<<polyse1, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = 'Log standard errors and relative standard errors for six of the polynomial models.'>>=
par(mfrow = c(4, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(ex_p2$se / ex_p2$estimate, main = '2nd Degree\nRSE')
plot(ex_p4$se / ex_p4$estimate, main = '4th Degree\nRSE')
plot(ex_p6$se / ex_p6$estimate, main = '6th Degree\nRSE')

@@

\section{Difficulties with Implementation}
